---
title: "Using the carrots data package"
author: "Peter Kamerman"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using the carrots data package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

The _carrots_ data package contains a reduced version of the carrots dataset found within the _lmerTest_ package. The reduced dataset includes data from all 103 consumers included in the original dataset, but preference scores are only reported for 6 of the 12 Danish carrot types. The other columns retained from the original dataset are consumer rating of carrot sweetness, bitterness and crispiness, the frequency at which each consumer consumed carrots, and the age, sex, income, and employment categories the consumers fell into.

To use the data, first install the package:

```{r install, eval = FALSE}
devtools::install_github('kamermanpr/carrots')
```

Once the package has been installed you can access the data as follows:

```{r access_wide, fig.align='center', fig.width=7, fig.height=5}
# Extract the wide format carrots data
wide <- carrots::carrots

# View the dataframe
head(wide)

# Examine the columns
str(wide)

# Generate a summary for each column
summary(wide)

# Plot relationship between preference and age group for Bolero carrots
boxplot(Preference ~ Age_group, 
        data = wide[wide$Product == 'Bolero', ],
        main = 'Bolero carrots: Preference rating by age group')
```

### Note

To access the carrots data with the attribute data (sweetness, bitterness, crispness) in long format, use: `carrots::long_carrots`.